---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)


  combined_candy_clean <- read_csv("../clean_data/combined_candy_clean.csv")
  candy_2015_clean <- read_csv("../clean_data/candy_2015_clean.csv")
  candy_2016_clean <- read_csv("../clean_data/candy_2016_clean.csv")
  candy_2017_clean <- read_csv("../clean_data/candy_2017_clean.csv")
```

## Question 1

What is the total number of candy ratings given across the three years. 
(Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
combined_candy_clean %>% 
  select(butterfinger:take_5) %>%
  pivot_longer(cols = butterfinger:take_5) %>% 
  summarise(total_num_of_ratings = sum(!is.na(value)))

# Utilised this code for validation that my pivot longer table worked.
# combined_candy_clean %>% 
#   relocate(gender,country,state_province_county, .before = butterfinger) %>% 
#   select(butterfinger:take_5) %>%
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   group_by(value) %>% 
#   count()
# 
#229481 + 264226 + 74013 = 567720
```

Answer: 
The total number of candy ratings across the three years was 567720. 



## Question 2

What was the average age of people who are going out trick or treating?

```{r}
combined_candy_clean %>% 
  select(age, trick_or_treating) %>% 
  filter(trick_or_treating %in% "Yes") %>% 
  summarise(trick_or_treaters_mean_age = round(mean(age, na.rm = TRUE )))

# validation
# combined_candy_clean %>% 
#   select(age, trick_or_treating) %>% 
#   group_by(trick_or_treating) %>% 
#   summarise(n())

```
The average age of people going trick or treating was 35 (rounded to the nearest
whole number). 


## Question 3
What was the average age of people who are not going trick or treating?

```{r}
combined_candy_clean %>% 
  select(age, trick_or_treating) %>% 
  filter(trick_or_treating %in% "No") %>% 
  summarise(non_trick_or_treaters_mean_age = round(mean(age, na.rm = TRUE )))
```
The average age of people going trick or treating was 39 (rounded to the nearest
whole number).



## Question 4
For each of joy, despair and meh, which candy bar received the most of these ratings? 

```{r}
 combined_candy_clean %>% 
  select(butterfinger:take_5) %>% 
  pivot_longer(cols = butterfinger:take_5) %>% 
  group_by(name) %>% 
  count(value) %>% 
  group_by(value) %>% 
  slice_max(n, n = 1, with_ties = FALSE)

#includes only candy items that I have defined as "candy bar"
combined_candy_clean %>% 
  select(butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, nestle_crunch, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         chick_o_sticks, three_musketeers, peanut_butter_bars, twix, coffee_crisp,
         dove_bars, mr_goodbar, whatchamacallit_bars, take_5) %>% 
  pivot_longer(cols = butterfinger:take_5) %>% 
  group_by(name) %>% 
  count(value) %>% 
  group_by(value) %>% 
  slice_max(n, n = 1, with_ties = FALSE)
  
```
 
 Answer:
 
 When we only look at items of candy which would be defined as a "candy bar",
 chick_o_sticks received the most despair ratings, any_full_sized_candy_bar had
 the most "JOY" ratings and  x100_grand_bar received the most "MEH" ratings. 

## Question 5
How many people rated Starburst as despair?

```{r}

combined_candy_clean %>% 
  filter(starburst == "DESPAIR") %>% 
  nrow()

```
Answer: 
1990 people gave Starburst a despair rating.

For the next three questions, count despair as -1, joy as +1, and meh as 0.

## Question 6
What was the most popular candy bar by this rating system for each gender in the 
dataset ?

```{r}
combined_candy_clean %>% 
  select(gender, butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         chick_o_sticks, three_musketeers, peanut_butter_bars, twix, coffee_crisp,
         dove_bars, mr_goodbar, whatchamacallit_bars, take_5) %>% 
  pivot_longer(cols = butterfinger:take_5) %>%
  mutate(value = case_when(value == "DESPAIR" ~ -1,
                           value == "MEH" ~ 0,
                           value == "JOY" ~ 1)) %>% 
  filter(!is.na(value)) %>% 
  group_by(gender, name) %>% 
  summarise(rating = sum(value), count = n()) %>% 
  arrange(gender, desc(rating), desc(count)) %>% 
  slice_max(n = 2, order_by = rating, with_ties = FALSE) %>% 
  select(gender, name, rating)

```
Answer: 
The most popular candy bar for each gender category was any_full_sized_candy_bar.

However if I remove any_full_sized_candy_bar from the list the most popular candy bar
by gender was  kit_kat for females, and  for males, I'd rather not say, and NA gender categories.
Twix was the favourite for the other gender category.

```{r}
# Please ignore this code chunk, I misread the question and did not assign the 
# values to the ratings. This code takes the candy with the highest number of 
# JOY ratings as being then organises by gender as required. I've left it in as 
# it's another way to interpret the popularity of a candy from the data set. 
# This code can be used to find the most popular candy on the list but is not 
# specific to items classified as "Candy Bars". 

# combined_candy_clean %>% 
#   select(gender, butterfinger:take_5) %>% 
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   filter(value == "JOY") %>% 
#   group_by(gender, name) %>% 
#   count(value) %>%
#   arrange(gender, desc(n)) %>% 
#   select(gender, name, n) %>% 
#   group_by(gender) %>% 
#   slice_max(n, n = 1, with_ties = FALSE)


# #This code includes columns with items that I have defined as a "candy bar"
# combined_candy_clean %>% 
#   select(gender, butterfinger, x100_grand_bar, any_full_sized_candy_bar,
#          caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
#          hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
#          mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
#          chick_o_sticks, three_musketeers, peanut_butter_bars, twix, coffee_crisp,
#          dove_bars, mr_goodbar, whatchamacallit_bars, take_5) %>% 
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   filter(value == "JOY") %>% 
#   group_by(gender, name) %>% 
#   count(value) %>%
#   arrange(gender, desc(n)) %>% 
#   select(gender, name, n) %>% 
#   group_by(gender) %>% 
#   slice_max(n, n = 1, with_ties = FALSE)
# 
# 
# #Below is the Candy bar analysis minus any_full_sized_candy_bar_column.
# 
# combined_candy_clean %>% 
#   select(gender, butterfinger, x100_grand_bar,
#          caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
#          hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
#          mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
#          chick_o_sticks, three_musketeers, peanut_butter_bars, twix, coffee_crisp,
#          dove_bars, mr_goodbar, whatchamacallit_bars, take_5) %>% 
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   filter(value == "JOY") %>% 
#   group_by(gender, name) %>% 
#   count(value) %>%
#   arrange(gender, desc(n)) %>% 
#   select(gender, name, n) %>% 
#   group_by(gender) %>% 
#   slice_max(n, n = 1, with_ties = FALSE)

```


## Question 7 
What was the most popular candy bar in each year?

```{r}

#2015 answer is any full sized candy bar for all candy items.
# candy_2015_clean %>% 
#   select(butterfinger:york_peppermint_patties) %>% 
#   pivot_longer(cols = butterfinger:york_peppermint_patties) %>% 
#   mutate(value = case_when(value == "DESPAIR" ~ -1,
#                            value == "MEH" ~ 0,
#                            value == "JOY" ~ 1)) %>% 
#   filter(!is.na(value)) %>%  
#   group_by(name) %>% 
#   summarise(rating = sum(value), count = n())  %>% 
#   arrange(desc(rating)) %>% 
#   head(5)
  

#2015 candy bars specific code
candy_2015_clean %>% 
  select(butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         three_musketeers, peanut_butter_bars, twix) %>% 
  pivot_longer(cols = butterfinger:twix) %>% 
  mutate(value = case_when(value == "DESPAIR" ~ -1,
                           value == "MEH" ~ 0,
                           value == "JOY" ~ 1)) %>% 
  filter(!is.na(value)) %>%  
  group_by(name) %>% 
  summarise(rating = sum(value), count = n())  %>% 
  arrange(desc(rating)) %>% 
  head(5)

#2016 answer is any full sized candy bar
# candy_2015_clean %>% 
#   select(butterfinger:york_peppermint_patties) %>% 
#   pivot_longer(cols = butterfinger:york_peppermint_patties) %>% 
#   mutate(value = case_when(value == "DESPAIR" ~ -1,
#                            value == "MEH" ~ 0,
#                            value == "JOY" ~ 1)) %>% 
#   filter(!is.na(value)) %>%  
#   group_by(name) %>% 
#   summarise(rating = sum(value), count = n())  %>% 
#   arrange(desc(rating)) %>% 
#   head(5)

#2016 candy bars specific code
candy_2016_clean %>% 
  select(butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         three_musketeers, twix, whatchamacallit_bars, mr_goodbar, dove_bars) %>% 
  pivot_longer(cols = butterfinger:whatchamacallit_bars) %>% 
  mutate(value = case_when(value == "DESPAIR" ~ -1,
                           value == "MEH" ~ 0,
                           value == "JOY" ~ 1)) %>% 
  filter(!is.na(value)) %>%  
  group_by(name) %>% 
  summarise(rating = sum(value), count = n())  %>% 
  arrange(desc(rating)) %>% 
  head(5)

#2017 answer is any full sized candy bar 

# candy_2017_clean %>% 
#   select(x100_grand_bar:york_peppermint_patties) %>% 
#   pivot_longer(cols = x100_grand_bar:york_peppermint_patties) %>% 
#   mutate(value = case_when(value == "DESPAIR" ~ -1,
#                            value == "MEH" ~ 0,
#                            value == "JOY" ~ 1)) %>% 
#   filter(!is.na(value)) %>%  
#   group_by(name) %>% 
#   summarise(rating = sum(value), count = n())  %>% 
#   arrange(desc(rating)) %>% 
#   head(5)

#2017 candy bars specific code
candy_2017_clean %>% 
  select(butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         three_musketeers, twix, whatchamacallit_bars, mr_goodbar, dove_bars, 
         coffee_crisp, take_5) %>% 
  pivot_longer(cols = butterfinger:whatchamacallit_bars) %>% 
  mutate(value = case_when(value == "DESPAIR" ~ -1,
                           value == "MEH" ~ 0,
                           value == "JOY" ~ 1)) %>% 
  filter(!is.na(value)) %>%  
  group_by(name) %>% 
  summarise(rating = sum(value), count = n())  %>% 
  arrange(desc(rating)) %>% 
  head(5)


```

## Question 8
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
# combined_candy_clean %>% 
#   select(country, butterfinger:take_5) %>% 
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   mutate(value = case_when(value == "DESPAIR" ~ -1,
#                            value == "MEH" ~ 0,
#                            value == "JOY" ~ 1)) %>% 
#   filter(!is.na(value)) %>%  
#   group_by(country, name) %>% 
#   summarise(rating = sum(value), count = n())  %>% 
#   arrange(desc(country), desc(rating)) %>% 
#   group_by(country) %>% 
#   slice_max(n = 1, order_by = rating, with_ties = FALSE)
  
  
  
  
  # filter(value == "JOY") %>% 
  # group_by(country, name) %>% 
  # count(value) %>%
  # arrange(country, desc(n)) %>% 
  # select(country, name, n) %>% 
  # group_by(country) %>% 
  # slice_max(n, n = 1, with_ties = TRUE)

# code below removes single values from analysis, this will remove some countries
# from table.
# combined_candy_clean %>% 
#   select(country, butterfinger:take_5) %>% 
#   pivot_longer(cols = butterfinger:take_5) %>% 
#   mutate(value = case_when(value == "DESPAIR" ~ -1,
#                            value == "MEH" ~ 0,
#                            value == "JOY" ~ 1)) %>% 
#   filter(!is.na(value)) %>%  
#   group_by(country, name) %>% 
#   summarise(rating = sum(value), count = n())  %>% 
#   arrange(desc(country), desc(rating)) %>% 
#   group_by(country) %>% 
#   slice_max(n = 1, order_by = rating, with_ties = FALSE)

#Code below is used for any item defined as a candy bar. 
combined_candy_clean %>% 
  select(country, butterfinger, x100_grand_bar, any_full_sized_candy_bar,
         caramellos, snickers, nestle_crunch, dark_chocolate_hershey, heath_bar,
         hershey_s_milk_chocolate, kinder_happy_hippo, kit_kat,
         mars, reggie_jackson_bar, milky_way, tolberone_something_or_other,
         chick_o_sticks, three_musketeers, peanut_butter_bars, twix, coffee_crisp,
         dove_bars, mr_goodbar, whatchamacallit_bars, take_5) %>% 
  pivot_longer(cols = butterfinger:take_5) %>% 
  mutate(value = case_when(value == "DESPAIR" ~ -1,
                           value == "MEH" ~ 0,
                           value == "JOY" ~ 1)) %>% 
  filter(!is.na(value)) %>%  
  group_by(country, name) %>% 
  summarise(rating = sum(value), count = n())  %>% 
  arrange(desc(country), desc(rating)) %>% 
  group_by(country) %>% 
  slice_max(n = 1, order_by = rating, with_ties = FALSE)

```

