---
title: "decathlon_data_analysis_questions"
author: "DC"
date: "2023-04-10"
output: html_document
---

# Analysis Questions

```{r}
library(tidyverse)

decathlon_info <- read_csv("clean_data/clean_decathlon.csv")
```
## Question 1. Who had the longest long jump seen in the data?
```{r}
decathlon_info %>% 
  select(name, long_jump_distance_m) %>% 
  arrange(desc(long_jump_distance_m)) %>% 
  head(1)
```

## Question 2. What was the average 100m time in each competition?
```{r}
decathlon_info %>% 
  group_by(competition) %>% 
  summarise(avg_100m_time_s = mean(`100m_time_s`))
  
```

## Question 3. Who had the highest total points across both competitions?

```{r}
decathlon_info %>% 
  select(name, total_points) %>% 
  arrange(desc(total_points)) %>% 
  head(1)
```
## Question 4. What was the shot-put scores for the top three competitors in each competition?

```{r}
decathlon_info %>% 
  group_by(competition) %>% 
  slice_min(n = 3, order_by = ranking) %>% 
  select(competition, ranking, shot_put_distance_m)
```



## Question 5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_info %>% 
  filter(`400m_time_s` < 50) %>% 
  summarise(avg_points_400m_time_less_than_50s = mean(total_points))

decathlon_info %>% 
  filter(`400m_time_s` > 50) %>% 
  summarise(avg_points_400m_time_more_than_50s = mean(total_points))

```
```{r}
decathlon_info %>% 
  group_by(`400m_time_s` < 50) %>% 
  summarise(mean(if(`400m_time_s` < 50 = TRUE)))
```


